---
title: "Homework 19 - 12 Oct 2022"
output: html_document
date: "2022-10-27"
---

This is the key to homework 19, which asked each of you to do some R work to revisit material from Chapters 2 and 3 of the GSWR book.

Let's set things up:
```{r}
rm(list = ls())
library(dplyr)
library(here)
library(ggplot2)
```

And pull in some data

```{r}
Leaves <- read.csv(here("Data", "Class_leaf_measurements.csv"))
```

Now use the `str()` function to examine the dataset

```{r}
str(Leaves)
```

Now let's take a look at column names
```{r}
names(Leaves)
```

Some of these column names are frustratingly non-informative. Let's fix them.

```{r}
names(Leaves)<- c("Tree_num", "Species", "Leaf_num", "DBH_cm", "petiole_length_cm", "total_length_cm", "middle_lobe_length_cm", "middle_lobe_width_cm", "person")
names(Leaves)
```

Much better!

Now convert species to a factor and determine the number of levels in the factor.

```{r}
Leaves$Species<-as.factor(Leaves$Species) #convert to factor data type
levels(Leaves$Species) #view levels of factor
```

We have many inaccuracies in how species names are recorded, so we need to fix them. We will do so with the `which()` function and indexing.

```{r}
Norway<-which(Leaves$Species == "Norway mapel" | Leaves$Species == "Norway Maple")
Leaves$Species[Norway] <- "Norway maple"
Red <- which (Leaves$Species == "red maple" | Leaves$Species == "Red Maple" | Leaves$Species == "red maple")
Leaves$Species[Red] <- "Red maple"
Sugar <- which (Leaves$Species == "sugar maple" | Leaves$Species == "Sugar Maple" )
Leaves$Species[Sugar] <- "Sugar maple"
Leaves$Species <- factor(Leaves$Species) #reset the factor levels
levels(Leaves$Species) #check that it worked
str(Leaves)
```

Now create a new dataset by subseting to get only the sugar maple data and so that you leave out the Person, Leaf number and tree number variables. Save the result as "Sugar"

```{r}
Sugar <- Leaves %>% filter(Species == "Sugar maple") %>% select(-person, -Leaf_num, -Tree_num)
```

Finally, you were asked to create a histogram of petiole lengths in the Sugar data set.

```{r}
ggplot(Sugar, aes(x = petiole_length_cm))+
     geom_histogram()
```







