---
title: "Homework 12 Key"
output: html_notebook
---
## Homework 12 Key

Make a conscious choice to detail your student/scientific thinking and to comment your code.


Add code to clear R's brain and to turn on the dplyr and here packages 

```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)
```

Import the csv file called “donnerville_percent_cover.csv” using stringsAsFactors = TRUE

Name the dataset “Cover” during import

```{r}
Cover <- read.csv(here("Data", "donnerville_percent_cover.csv"), stringsAsFactors = TRUE)
```

Examine the data set using at least two different functions.
```{r}
glimpse(Cover)
head(Cover)
str(Cover)
tail(Cover)
names(Cover)
```

Consider renaming column headings; this choice is up to you.

Let's do it to see how it works

```{r}
Cover <- Cover %>% rename(Plot_type = Plot.type, Plot_dim = Plot.dim)
names(Cover)

names(Cover)<- c("Site", "Date", "Plot_type",
 "Plot_dim" , "Slope"  ,   "Aspect",   
"Plot"  ,    "Type"   ,   "Percent"  )

names(Cover)
```

Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1.

```{r}
Cover %>% filter(Site == "Donnerville 1") %>% group_by(Type) %>% summarise(
     avg_percent_cover = mean(Percent)
)
```

Determine the average percent vascular plant cover across all sites.

```{r}
PerCov <- Cover %>% filter(Type == "Vascular") %>% summarise(
     avg_percent = mean(Percent)
)
```

