---
title: "Homework 23 Key"
author: "Erika Barthelmess"
date: "`r Sys.Date()`"
output: html_document
---

Beginning Data Analysis - Data Visualization
Using the soil data provided in today's course module, create a scatterplot, box plot and histogram for the SOM_percent column (which represents the % soil organic matter by weight in the soil sample).  Then create the same three plots for Ca ppm (parts per million soil calcium).  

Construct these plots in a Homework23.Rmd file and push the changes to your GitHub repo. Send me a message in slack when you are finished.

Set things up
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)

soil<-read.csv(here("Data", "soil_data.csv"))
```
First, in order to create boxplots and compare them between forests, we need a forest variable. We're going to have to extract it from Field.Name

```{r}
soil$Field.Name<-as.factor(soil$Field.Name)
levels(soil$Field.Name)
```
We can do this most easily with string matching
```{r}
library(stringr)
DG<-str_detect(soil$Field.Name, "DEGRASSE")
DGfix<-which(DG == TRUE)
soil$Forest<-"forest"
soil$Forest[DGfix]<-"Degrasse"

DON<-str_detect(soil$Field.Name, "DONNERVILLE")
DONfix<-which(DON == TRUE)
soil$Forest[DONfix]<-"Donnerville"

KT<-str_detect(soil$Field.Name, "KIP TRACT")
KTfix<-which(KT == TRUE)
soil$Forest[KTfix]<-"Kip Tract"

PEA<-str_detect(soil$Field.Name, "PEAVINE")
PEAfix<-which(PEA == TRUE)
soil$Forest[PEAfix]<-"Peavine"

SH<-str_detect(soil$Field.Name, "SOUTH HAMMOND")
SHfix<-which(SH == TRUE)
soil$Forest[SHfix]<-"South Hammond"

soil$Forest<-as.factor(soil$Forest)
levels(soil$Forest)
```
Create a scatterplot, box plot and histogram for the SOM_percent column

Let's start with a histogram
```{r}
ggplot(soil, aes(x = Organic.Matter.. ))+
     geom_histogram()
```
Now let's go for a boxplot
```{r}
ggplot(soil, aes(Forest,Organic.Matter.. ))+
     geom_boxplot()
```
And now a scatter plot
```{r}
ggplot(soil, aes(x = Forest, y = Organic.Matter..))+
     geom_point()
```
Now create same three plots for  Ca ppm (parts per million soil calcium). 

```{r}
ggplot(soil, aes(x = Ca.ppm))+
     geom_histogram()
```
Scatterplot
```{r}
ggplot(soil, aes(Forest, Ca.ppm))+
     geom_point()
```
Box plot
```{r}
ggplot(soil, aes(Forest, Ca.ppm))+
     geom_boxplot()
```


