---
title: "Biostats - Homework 21"
output: html_notebook
---

The goal of this homework was twofold:
1. Practice using a github workflow
2. Begin learning to make plots in RStudio

From the assignment on canvas:

For today's assignment, please do the following:

- In your homework project on RStudio server (the above-mentioned one that is linked to github), create a new .Rmd file in your Code folder called Homework-21.Rmd

- Go through Chapter 4 of Getting Started With R and write out all the code from the chapter in your .Rmd file.

- You will need the compensation dataset (used in prior GSWR homeworks) to complete this exercise. Place a copy of the compensation dataset in the Data folder for your project and point to it when you load your dataframe.

- When you are finished, commit and push your changes to github and then send me a message in the #homework channel on slack that includes the link to your github homework repository.

- Pat yourself on the back, grab a preferred beverage, and feel all-powerful! You just crushed an important open science skill by sharing your code on github!

Let's set things up:
```{r}
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)

compensation <- read.csv(here("Data", "compensation.csv"))
```
## Data Visualization (the theme of this chapter)

### Create a scatterplot
We are going to work on three plot types, the first of which is the scatterplot:

```{r}
ggplot(compensation, aes(x = Root, y = Fruit))+
     geom_point()
```
















