---
title: "Homework 26"
author: "Erika"
date: "11/2/2022"
output: html_document
---
Set things up
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Now let's pull in the data:
```{r}
soil <- read.csv(here("Data", "soil_reduced-1.csv"),stringsAsFactors = TRUE)
```
Before we begin any analysis, we need to check our data for possible errors/outliers. Let's begin with that.

A. Forest names
```{r}
levels(soil$Forest)
```
B. Let's check soil organic matter and soil calcium
STart with a histogram for SOM
```{r}
ggplot(soil, aes(x = SOM_percent))+
     geom_histogram()
```

Now check soil calcium
```{r}
ggplot(soil, aes(x = Ca_ppm))+
     geom_histogram()
```

Let's also just run a summary on all the variables:
```{r}
summary(soil)
```

## We'd like to ask the following questions from these data:

### 1. Is there a relationship between Soil organic matter (SOM, %) and soil calcium (Ca_ppm)?

Both variables are continuous
Let's make a scatterplot
#### Plot relationship
```{r}
ggplot(soil, aes(x = SOM_percent,  y = Ca_ppm, color = Forest))+
     geom_point()+
     theme_bw()
```
From looking at the data, it appears there may be a weak positive relationship b/w SOM and Ca

Can I guess the slope or y-intercept?
Slope c. 1000/53 = 18ish, y-int c. -5? 0? right around there.

What statistical model fits these data?
Both response and predictor are continuous so linear regression seems appropriate.

So, let's construct the model:
```{r}
SOM_CA_mod <- lm(Ca_ppm ~ SOM_percent, data = soil)
```
Now I need to check model assumptions before interpreting.

```{r}
autoplot(SOM_CA_mod)
```
Based on my observations of the autoplot results, I'm modestly comfortable that my data meet the lm() assumptions.

Now look at model outcome
```{r}
anova(SOM_CA_mod)
```
```{r}
summary(SOM_CA_mod)
```
























2. Does the abundance of worms differ between Donnerville and the Kip Tract?

3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?

4. Does the abundance of zinc (Zn_ppm) differ among forests?