---
title: "Homework 14 - Using dplyr to fix mistakes"
output: html_notebook
---

Today's homework is designed to help you reinforce concepts from GSWR Chapter 3 and to push you to apply functions from the dplyr package more independently. You will use the attached data set about testing two drugs for effectiveness as antibiotics.

Set up your session by writing a code block with the "usual" material near the top of your .Rmd file.
Import the data using read.csv but without stringsAsFactors = TRUE and name the data frame "DF"  

Set things up and bring in the data
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(here)

DF <- read.csv(here("Data", "Homework Data", "drug_resistance.csv"))
```

Run a summary of DF and see what it tells you.

```{r}
summary(DF)
```
Now import the data again, still using read.csv but this time adding stringsAsFactors = TRUE. Again name the dataframe DF. It will write over the first data frame you imported. Run a summary of DF and see how it differs from the first time you ran it.

```{r}
DF <- read.csv(here("Data", "Homework Data", "drug_resistance.csv"), stringsAsFactors = TRUE)
summary(DF)
```

Make some notes in your .Rmd file about how the summary() command reacts to data that do vs. do not have explicit factor variables in the data set.

When the data are structured upon import to have factors, the summary function is able to provide more information about those variables than when they are imported as character data.

Now use either the function unique() or levels() to examine each of the text-based columns of data. In which variables do you see errors?

```{r}
unique(DF$Treatment)
unique(DF$tech.name)
```
Errors are present in:

- Treatment. It shows 9 levels but should only have 3 (control, drug x 2 and drug x 3)

- tech.name has errors. Brian and Becca are spelled multiple ways. There should just be: jose, brian, mary and becca.

Using the techniques we discussed in class, repair the data set so that there are the appropriate number of choices for each of the character/factor variables.

Let's start with Treatment. 

```{r}
unique(DF$Treatment)
```
Find all the cases of Control and cantrol and make them all control

```{r}
fix <- which(DF$Treatment == "Control"|DF$Treatment == "cantrol")
DF$Treatment[fix]<- "control"
unique(DF$Treatment)
```
Now work on the drug treatments
```{r}
fix <- which(DF$Treatment == "Drugx2"| DF$Treatment == "drug2")
DF$Treatment[fix] <- "drug x 2"
unique(DF$Treatment)

fix <- which(DF$Treatment == "drug3" | DF$Treatment == "Drug x 3")
DF$Treatment[fix] <- "drug x 3"
unique(DF$Treatment)
```
Notice that R still thinks there are 9 levels. We need to relevel them by making Treatment a factor variable again.
```{r}
DF$Treatment <-factor(DF$Treatment)
levels(DF$Treatment)
```
Now use similar steps to fix the tech names
```{r}
unique(DF$tech.name)
```

```{r}
fix <- which(DF$tech.name == "Brian" | DF$tech.name == "brain")

DF$tech.name[fix]<-"brian"

fix <- which(DF$tech.name == "Becca")
DF$tech.name[fix] <- "becca"

DF$tech.name <- factor(DF$tech.name)
levels(DF$tech.name)
```





