---
title: "Homework 26"
author: "Erika"
date: "11/2/2022"
output: html_document
---
Set things up
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Now let's pull in the data:
```{r}
soil <- read.csv(here("Data", "soil_reduced-1.csv"),stringsAsFactors = TRUE)
```
Before we begin any analysis, we need to check our data for possible errors/outliers. Let's begin with that.

A. Forest names
```{r}
levels(soil$Forest)
```
B. Let's check soil organic matter and soil calcium
STart with a histogram for SOM
```{r}
ggplot(soil, aes(x = SOM_percent))+
     geom_histogram()
```

Now check soil calcium
```{r}
ggplot(soil, aes(x = Ca_ppm))+
     geom_histogram()
```

Let's also just run a summary on all the variables:
```{r}
summary(soil)
```

## We'd like to ask the following questions from these data:

### 1. Is there a relationship between Soil organic matter (SOM, %) and soil calcium (Ca_ppm)?

Both variables are continuous
Let's make a scatterplot
#### Plot relationship
```{r}
ggplot(soil, aes(x = SOM_percent,  y = Ca_ppm, color = Forest))+
     geom_point()+
     theme_bw()
```
From looking at the data, it appears there may be a weak positive relationship b/w SOM and Ca

Can I guess the slope or y-intercept?
Slope c. 1000/53 = 18ish, y-int c. -5? 0? right around there.

What statistical model fits these data?
Both response and predictor are continuous so linear regression seems appropriate.

So, let's construct the model:
```{r}
SOM_CA_mod <- lm(Ca_ppm ~ SOM_percent, data = soil)
```
Now I need to check model assumptions before interpreting.

```{r}
autoplot(SOM_CA_mod)
```
Based on my observations of the autoplot results, I'm modestly comfortable that my data meet the lm() assumptions.

Now look at model outcome
```{r}
anova(SOM_CA_mod)
```
```{r}
summary(SOM_CA_mod)
```
That is how far we got in class. Now we need to make a plot that shows the trend line.

Now let's make a figure that shows off our relationship in light of our statistical results.

```{r}
ggplot(soil, aes(SOM_percent, Ca_ppm))+
     geom_point()+
     geom_smooth(method = "lm")+
     ylab("Calcium (parts per million)")+
     theme_bw()
```
This explains why my original estimate for slope and y-intercept was wrong - when I eyeballed the figure, I thought the line would be steeper.


### 2. Does the abundance of worms differ between Donnerville and the Kip Tract?

Now let's go through the process again.

We'll filter the data down to just Donnerville and Kip Tract, but before we do so, we need to make sure that the forest names are ok so that we don't miss anything.
```{r}
levels(soil$Forest)
```
Looks good. Now let's filter to just the data we need.

```{r}
worms<-soil %>% filter(Forest == "DONNERVILLE" | Forest == "KIP TRACT")

worms$Forest<- factor(worms$Forest)
```
Now we'll plot the worm data to look for outliers and also run the `summary()` function..

```{r}
ggplot(worms, aes(Worms))+
     geom_histogram()

summary(worms$Worms)
```
These data look reasonable. So we're ok as far as error checking goes.

Now we'd like to plot our data and visualize the relationship.  We need a group by and summarise first.

```{r}
totals <- worms %>% group_by(Forest) %>% summarise(total.worms = sum(Worms))
```

Because we have count data, this is a time when using a bar graph is ok.

```{r}
ggplot(totals, aes(x = Forest, y = total.worms))+
     geom_bar(stat = "identity", position = "dodge")
```
It seems like there are more worms on the Kip Tract. It is a much more disturbed forest, so that isn't surprising.

Now we need to think about the structure of our data. Forest is a categorical variable and the number of worms is a discrete variable, so the likely test is a Chi-square test for independence. In order to run this test, we need our data summarized in a matrix. Let's do that:

We will use the function `xtabs()` to do so:

```{r}
worms.mat <- xtabs(Worms ~ Forest, data = worms)
worms.mat
```

Now we can run the test
```{r}
chisq.test(worms.mat)
```
Interpretation
Our null hypothesis is that there is no difference in worm abundance between the Kip Tract and Donnerville. However, these is only a tiny chance (0.0078) that we would see this difference in worm #s by chance alone, suggesting that infact we can reject our null and conclude that there are signifcantly more worms on the Kip Tract than at Donnerville.


### 3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?
Ok - you know the drill. We determined above that there was no problem with the Forest column so we can use it to subset the data to our focal forests.

```{r}
chem_data<-soil %>% filter(Forest == "PEAVINE" | Forest == "SOUTH HAMMOND")

chem_data$Forest <- factor(chem_data$Forest)
```

Now let's plot the data. This time, how about a boxplot.

```{r}
ggplot(chem_data, aes(Forest, Mg_ppm))+
     geom_boxplot()+
     geom_point(alpha = 0.5)
```
It is possible that South Hammond has an outlier. Take a look
```{r}
chem_data$Mg_ppm
```
121 seems a bit high, but there is a 92, and quite a bit of range, so I'll keep it.

Based on this view, it does appear that I could run the test. Since I have a factor variable (x) and a continuous variable (y) and since there are only two groups in my x variable, this should be a 2-sample t-test. Let's work with that.

Begin by plotting the data

```{r}
ggplot(chem_data, aes(x = Mg_ppm))+
     geom_histogram(binwidth=10)+
     facet_wrap(~Forest, ncol = 1)+
     theme_bw()
```

They look fairly similar, though the distribution skews far enought to the right that there may be a difference.

Let's run the model

Standard error is standard deviation  divided by square root of sample size
```{r}
worm_means<-chem_data %>% group_by(Forest) %>% summarise(
     mean_worms = mean(Worms,na.rm = TRUE), 
     sem_worms = (sd(Worms, na.rm = TRUE)/sqrt(length(Worms)))
     
)
```

Now let's do the t-test
```{r}
t.test(Worms ~Forest, data = chem_data)
```

We cannot reject the null hypothesis. Now for a better figure:

```{r}
ggplot(worms, aes(Forest, Worms))+
     geom_boxplot()+
     geom_point(position = "jitter", size = 2, alpha = 0.5)+
     labs(y = "Number of worms")+
     theme_bw()
```

### 4. Does the abundance of zinc (Zn_ppm) differ among forests?
Ok. Last one. Let's step through our process again.

Begin by checking the data. We've already looked at forests. How about zinc?

```{r}
summary(soil$Zn_ppm)
```

Looks like there could be an outlier. Let's view a histogram.
```{r}
ggplot(soil, aes(x = Zn_ppm))+
     geom_histogram()
```
It looks reasonable that the point with value 22 might be just the tail of real values. We'll keep it.

Now that I know my data are ok, I should plot the data. My x variable is categorical, y is continuous. A boxplot would be a good choice here.

```{r}
ggplot(soil, aes(Forest, Zn_ppm))+
     geom_boxplot()+
     geom_point(size = 2, alpha = 0.5, position = "jitter")
```
It looks like the zinc levels in Degrasse are pretty low but that there is a lot of spread at the other sites. So I'm not confident that there is enough between group variance relative to within group to yield a statistically significant difference overall.

Time for our test

This looks like a one-way ANOVA because there are more than 2 levels of my categorical variable and because my response variable is continuous.

Build the model
```{r}
Zn_mod <- lm(Zn_ppm ~ Forest, data = soil)
```
Now we need to check the assumptions 

```{r}
autoplot(Zn_mod, smooth.color = NA)
```
Looks reasonable.

So now let's look at the results of our model and do some interpretation.

```{r}
anova(Zn_mod)
```
From this table, we see that there is not a significant difference in mean Zinc levels among forests. The F value, which represents the ratio of between group to within group variation, is small and has a correspondingly small p-value.

Let's look at the summary

```{r}
summary(Zn_mod)
```

